---
title: "HW4"
author: "Jillian Sarazen"
date: "2/5/2020"
output: html_document
---



#### 1. Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable zand print the value stored in z.
##### a)

```{r}
x <- 1.1
a <- 2.2
b <- 3.3
z <- (x^a^b)
print(z)
```

##### b)
```{r}
z <- ((x^a)^b)
print(z)
```


##### c)
```{r}
z <- 3*(x^3) + 2*(x^2) + 1
print(z)
```


#### 2. Using the rep and seq functions, create the following vectors:
##### a)
```{r}
z <- c(seq(1,8,1),seq(7,1, -1))
paste(z)
```
###### use c() function to combine two sequences

##### b)
```{r}
z <- c(1, rep(2,2), rep(3,3), rep(4,4), rep(5,5))
paste(z)
```


##### c)
```{r}
z <- c(5, rep(4,2), rep(3,3), rep(2,4), rep(1,5))
paste(z)
```


#### 3. Create a vector of two random uniform numbers...

```{r}
z <- runif(2)

r <- sqrt(z[1]^2) + sqrt(z[2]^2)
t <- atan(z[2]/z[1])

j <- c(r,t)
paste(j)
```
##### check if this works?

#### 4 Suppose that queue <- c("sheep", "fox", "owl", "ant") and that queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update the queue successively as:
```{r}
q <- c("sheep", "fox", "owl", "ant")
paste(q)

q <- append(q,"serpent")
q <- q[-(1)]
# this takes the first element
q <- append(q, "donkey", 0)
# append function to add an element in the vector, 0 puts it after the 0 element, aka first
q <- q[-(5)]
# take out serpent
q <- q[-(3)]
# take out owl
q <- append(q, "aphid", 2)
# add in aphid before the ant
which(q=="aphid")
# extracts a value when it is equal to another value

paste(q)
```
#### 5 Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.

```{r}
# make a sequence of 100 variables
a <- seq(1,100,1)
paste(a)

# this takes out all values from the vector that are divisable by 2, 3, and 7. this works! [] remove those values from the vector. %% is an operator, still not totally sure what it means though. != is the function for not equal
b <- a[a%%2!=0 & a%%3!=0 & a%%7!=0]
paste(b)

```

