---
title: "HW 5"
author: "Jillian Sarazen"
date: "2/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1) Assign to the variable n_dims a single random integer between 3 and 10.

```{r}
# sample function, integers between 3 and 10, choose 1 number
n_dims <- sample(3:10,1)
print(n_dims)
```

- Create a vector of consecutive integers from 1 to n_dims2.
```{r}
n_dims2 <- seq(1:n_dims^2)
print(n_dims2)
```

- Use the sample function to randomly reshuffle these values.

```{r}
# just sample to reshuffle the vector I made above
sample(n_dims2)
```

- create a square matrix with these elements.
```{r}
m <- matrix(data=n_dims2,nrow=n_dims,ncol=n_dims)
```

- print out the matrix.
```{r}
print(m)
```

- find a function in r to transpose the matrix.

```{r}
# t() is the transpose function in 2, which switches columns and rows
m2 <- t(m)
```

- print it out again and note how it has changed.

```{r}
print(m2)
```

- calculate the sum and the mean of the elements in the first row and the last row.

```{r}
# function is the first element, then list the matrix name, then call out [row,column] using [ ]
m_sum <- sum(m2[1,])
print(m_sum)
m_mean <- mean(m2[5,])
print(m_mean)
```

- read about the eigen() function and use it on your matrix

```{r}
ev <- eigen(m2)
print(ev)
```

- look carefully at the elements of $values and $vectors. What kind of numbers are these?


#### not sure what type of numbers they are...


- dig in with the typeof() function to figure out their type.

```{r}
typeof(ev$values)
typeof(ev$vectors)
```

- if have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.


##### yes this works, used the n_dims variable throughout the code so that it can be customized 


2) Create a list with the following named elements:

- my_matrix, which is a 4 x 4 matrix filled with random uniform values

- my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.

- my_letters, which is a 26-element vector of all the lower-case letters in random order.



```{r}
# first I made the second logical statement at the beginning. Set up a vector of 100 random numbers of either 0 or 1, and then the replace function would say if its a true or false
log <- sample(0:1,100,replace = TRUE)

# this list has all 3 elements, a matrix of 16 random uniform elements in 4x4. Then the logical version of the log vector I made above. Then in the letters vector is a randomly ordered vector of lowercase letters
my_list <- list(my_matrix = matrix(data=runif(16),nrow=4,ncol=4),
                my_logical = as.logical(log),
                my_letters = sample(letters))

str(my_list)
```

- use the typeof() function to confirm the underlying data types of each component in this list
```{r}
# this didn't work in one statement, but this pulls out the data type for each element in the list

typeof(my_list[[1]])
typeof(my_list[[2]])
typeof(my_list[[3]])
```


-create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.


```{r}
## [[ ]] selects the element in the list

my_list2 <- list(my_list[[1]][2,2],
                 my_list[[2]][2],
                 my_list[[3]][2])
```

- combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r}

vec_1 <- c(my_list2)

```


- what is the data type of this vector?

```{r}
typeof(vec_1)
# not sure if this data type should be a list?
```


3) Create a data frame with two variables (= columns) and 26 cases (= rows).
- call the first variable my_unis and fill it with a random uniform value from 0 to 10
- call the second variable my_letters and fill it with 26 capital letters in random order.

```{r}

my_units <- runif(26, min = 0, max = 10)
my_letters <- sample(LETTERS)
new_data1 <- data.frame(my_units,my_letters, stringsAsFactors = FALSE)
print(new_data1)

```


- for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.

```{r}
# for this use brackets and the sample function. assign NA to those 4 values

new_data1$my_units[sample(1:26, size = 4)] <- NA 

```

- for the first variable, write a single line of R code to identify which rows have the missing values.

```{r}
new_data_missing <- which(is.na(new_data1$my_units))
print(new_data_missing)
```


- for the second variable, sort it in alphabetical order

```{r}

sort(new_data1$my_letters)

```



- calculate the column mean for the first variable.

```{r}
# since the my units column was changed above with the NA replacement, need to run the first part of the code with the data frame designation again before running this

my_units <- runif(26, min = 0, max = 10)
my_letters <- sample(LETTERS)
new_data1 <- data.frame(my_units,my_letters, stringsAsFactors = FALSE)
print(new_data1)

mean_my_units <- mean(new_data1$my_units)
print(mean_my_units)

```

