---
title: "HW_7"
author: "Jillian Sarazen"
date: "2/26/2020"
output: html_document
---

## Homework #7: Creating Fake Data Sets To Explore Hypotheses

#### Data in each of your treatment groups follow a normal distribution. Specify the sample sizes, means, and variances for each group that would be reasonable if your hypothesis were true.

```{r}

# this section of code sets the parameters of the data, SD, mean, sample size
n_group <- 2 # number of treatment groups
n_names <- c("WCB1", "WCB2") # WCB1 = woodchip bioreactor 1 (and 2, there are 2 outflows)
n_size <- c(14,15) # size is number of samples, aka number of storm events
n_mean <- c(23.79,45.39) # means of the two groups
n_sd <- c(13.57,25.24)
t_group <- rep(n_names,n_size) # this makes a t_group that has names of the treatment groups repeated for the whole sample size
t_group
table(t_group)

```


# this chunk of code sets up the data frame, one column is id, then t_group (made above, all the treatments)
```{r}

id <- 1:(sum(n_size)) # the sum of both samples from both groups
# random normal numbers to make up the value column (in my case TN values) 
res_var <- c(rnorm(n=n_size[1],mean=n_mean[1],sd=n_sd[1]), 
             rnorm(n=n_size[2],mean=n_mean[2],sd=n_sd[2]))

ano_data <- data.frame(id,t_group,res_var) 
str(ano_data)

```





## Running ANOVA model

```{r}

# ~ means "as a function of"
ano_model <- aov(res_var~t_group,data=ano_data)
print(ano_model) # to get the model
print(summary(ano_model)) # gets all elements of the model 
z <- summary(ano_model) # save as a vector
str(z)
aggregate(res_var~t_group,data=ano_data,FUN=mean) # separates out by the t_group and gives the mean of each group. applying a common operation to several columns

unlist(z)
unlist(z)[7] # takes the f value out, need to go in the summary section and figure out what you want and the number
ano_sum <- list(Fval=unlist(z)[7],
                probF=unlist(z)[9])
ano_sum

```

## Plot data using a boxplot

```{r}

# ggplot for ANOVA data ---------------------------------------------------
library(ggplot2)


ano_plot <- ggplot(data=ano_data,
                   aes(x=t_group, # aes=specify asthetics
                       y=res_var,
                       fill=t_group)) +# fill to specify a grouping variable
                      geom_boxplot()
print(ano_plot)

```


# My observations about from rerunning the code:

The first time that I ran the code I got a significant P value, but the next 2 times the P value was not significant. 

#### Change the parameters around: Made sample size to 10 compared to 15.
Sample size is important to see a difference, because when I had parameters of the same SD, sample size of 30 (increased from 15), and means of 25 and 30. I got a significant difference. Change the sample sizes down to 10, and the P value is now 0.1466. 

#### Ran this code with my real data for sd, mean, and sample size (however they are not actually normally distributed)
I got a P value of 0.002, which makese sense that the two groups (outflows) seem to be visually different from each other. 


#### Alternatively, for the effect sizes you originally hypothesized, what is the minimum sample size you would need in order to detect a statistically significant effect? Again, run the model a few times with the same parameter set to get a feeling for the effect of random variation in the data.
Even for a sample size of 6, I still got a significant P-val

